---
title: 'Lab Week 5: Times Series and Map'
author: "Anna Abelman"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

#### Attach packages
```{r}
# For general stuff:
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)

# For ts stuff: 
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)

# For spatial stuff: 
library(sf)
library(tmap)
library(mapview)

```

#### Monthly US energy consumption (renewables)
```{r}
us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()
```

 - Make description all lowercase
 - Only keep observation where "description" variable contains "comsumption"
 - Remove any observations where "description" variables contains "total"

```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% #lowercase version
  filter(str_detect(description, pattern = "consumption")) %>%
  filter(!str_detect(description, pattern = "total")) #exclamation points means to exclude that function
  
```

#### Convert `yyyymm` column to a date
```{r}
renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% #creates a new column that only contains month and year
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

#make a version where I have the month and year in separate columns
renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% #label = true makes it the month name and not the number
  mutate(year = year(yr_mo_day))
```

#### Look at the data
```{r}
renew_gg <- ggplot(data = renew_date, aes(x = month_sep,
                                          y = value, 
                                          group = description))+
  geom_line(aes(color = description))+ #if adding the color = , the group function is no longer necessary
  theme_minimal()

#View(palettes_d_names) to see the different colors palettes from the paletteers function
  
renew_gg
```

Updating my colors with paletteer palettes:
```{r}
renew_gg +
  scale_color_paletteer_d("tvthemes::rickAndMorty") #d because using discrete
```

#### Coerce renew_parsed to a tsibble

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```

Let's look at our ts data in a couple different ways:
```{r}
renew_ts %>% autoplot(value)

renew_ts %>% gg_subseries(value)

renew_ts %>% gg_season(value)

#or use ggplot

#ggplot(data = renew_parsed, aes(x = month, y = value, group = year))+
#geom_line(aes(color = year))+
#facet_wrap(~ description, ncol = 1, scale = "free", strip.position = right)
```

#### Just look at the hydroelectric consumption

```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
hydro_ts %>% gg_season(value)
```

#### What if I want quarterly average consumption for hydro?

```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarize(
    avg_consumption = mean(value)
  )
```

#### Decompose the hydro_ts

```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))

components(dcmp) %>% autoplot()

```

```{r}
hist(components(dcmp)$remainder) #looks somewhat normally distributed
```

Now look at the ACF:
```{r}
hydro_ts %>% 
  ACF(value) %>% 
  autoplot()
```

DANGER DANGER!
```{r}
hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value)
  ) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010))
```

#### Make a world map

```{r}
world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")
mapview(world)
```




















